---
title: "Ames Housing price analysis"
description: |
  This blog post includes EDA, Modeling and how to improve model performance and how Model Assessment and Diagnostics is done. Here, I am not predicting any price, in this blog you just get to know how to do analysis of price of the house with the help of single linear model.
author:
  - name: Jaykumar A Patel
date: 2021-05-04
preview: housing.png
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
---

# Problem Statment      
- Analyze the features affecting the price of the house and understand the linear equation for the feature. Understand how we can interpret the linear equation and with the help of that equation we can predict the price.


# Univariate analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(kableExtra)
```

```{r, include=FALSE}

knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.align = "center")

```

```{r packages, warning = FALSE, message = FALSE}

library(scales)
library(lares)
library(tidyverse)
library(modelsummary)
library(ggthemes)
library(gridExtra)
library(broom)

```

```{r, include=FALSE}

themes <-  function()
  {
  theme(
        plot.title = element_text(color = "#651a34", size = 20, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, color = "#845EC2",  face = "italic" ),
        axis.title.x = element_text(color = "#651a34", face = "bold",margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(color = "#651a34", face = "bold", margin = margin(t = 0, r = 10, b = 0, l = 0)),
        text = element_text(colour = "#BCF7E4"),
        legend.background = element_rect(fill="#B0E0E0",size=0.5, linetype="solid"),
        legend.title = element_text(colour="black", size=10, face="bold"),
        legend.text = element_text(color = "#545479"),
        legend.position = "none",
        legend.direction = "horizontal" 
        )
}

colors <- c('#845EC2', '#D65DB1', '#FF6F91', '#651a34', "#FFC75F", "#A8F78B","#9367C4", "#4B4453")

```

```{r}

data(ames, package = "modeldata")

ames$Garage_Cars <- factor(ames$Garage_Cars)

```

```{r}
num <- ames %>%
  select(where(is.numeric))
```

## Exploring the univariate distribution of the response variable Sale price.

```{r}
ames %>% 
  ggplot(aes(Sale_Price))+    
  geom_histogram(aes(y = stat(density)), color = "#13B4FA",fill = "#FF6F91") +
    geom_density(fill = "#845EC2", alpha = 0.5, color = NaN)+
    geom_vline(xintercept = mean(ames$Sale_Price), lwd = 0.8, linetype = "dashed", color = "#845EC2" )+
     geom_label(label= paste("Mean", round(mean(ames$Sale_Price), 2), sep =": "), x=250000, y=0.000007, label.padding = unit(0.55, "lines"), label.size = 0.25, fill = "#845EC2", color = "white"
  )+
    geom_vline(xintercept = median(ames$Sale_Price), lwd = 0.8, linetype="twodash", color = "#4B4453")+
     geom_label(label= paste("Median", round(median(ames$Sale_Price), 2), sep =": "), x=100000, y=0.000006, label.padding = unit(0.55, "lines"), label.size = 0.25, fill = "#4B4453", color = "white"
  )+
    labs(title = "Distribution of Sales Price",
         x = "Total Price",
         y = "density") +
  scale_x_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1))+
    theme_replace()+
    themes()
```

-   As we look at the mean and median, `median < mean` so we can say that our variable is right skewed. But it is not much skewed so we can consider it as normally distributed because in real world we never get perfectly distributed data.\
-   The reason behind this skewness can be outliers present in the response variable..

## Find outlier(s) in your original response variable

```{r}
ames %>% 
  ggplot(aes(y= Sale_Price))+
  geom_boxplot(outlier.color = "#13B4FA" )+
  labs(title = "Outliers in Sale Price",
       y = "Sale Price") +
  theme_replace()+
  scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1))+
  themes()

```

-   We have found the we have some outliers present in the data which can be reason for the skewness.

# Bivariate analysis


-   To do the bivariate analysis we need to find the features which are highly correlated to the response variable that is house price.

## Heat Map to find highly correlated features.

```{r}

corr_var(num, # name of dataset
  Sale_Price, # name of variable to focus on
  top = 3 # display top 5 correlations
) 
  
```

We can see top 3 numerical feature that are directly correlated to the response variable.

1.  **Gr_Liv_Area** Above Ground Living Area in sqft
2.  **Garage_Area** Garage Area in sqft
3.  **Total_Bsmt_SF** Total basement Size in sqft

Now, We look at each feature one by one and see if we can improve the relation?

## Relation between Sale Price and Above ground living area size

```{r}

data1 <- ames %>% 
          filter((!Gr_Liv_Area>3000)) %>% 
          select(Gr_Liv_Area,Sale_Price)

Gr_lv1 <- data1 %>% 
  ggplot(aes(Gr_Liv_Area, Sale_Price)) +
  geom_point() +
  labs(title = "Relation without outliers",
         x = "Above ground living area size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

Gr_lv2<- ames %>% 
  ggplot(aes(Gr_Liv_Area, Sale_Price)) +
  geom_point() +
  labs(title = "Relation with outliers",
         x = "Above ground living area size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

  
grid.arrange(Gr_lv2, Gr_lv1, ncol=2)
```

-   Here, we can say that we have some positive relation between sale price and Above Ground Living Area.

    -   But we can see there are some outliers in Above Ground Living Area, which leads to the less weak relation.

-   The Correlation coefficient before removing outliers is **`r cor(ames['Gr_Liv_Area'], ames['Sale_Price'])`**

-   The Correlation coefficient after removing outliers is **`r cor(data1['Gr_Liv_Area'], data1['Sale_Price'])`**

## Relation between Sale Price and Size of garage

```{r}

data2 <- ames %>% 
          filter((!Garage_Area>1250) & (Garage_Area > 0)) %>% 
          select(Garage_Area,Sale_Price)

Gar1 <- data2 %>% 
  ggplot(aes(Garage_Area, Sale_Price)) +
  geom_point() +
  labs(title = "Relation without outliers",
         x = "Size of garage",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

Gar2 <- ames %>% 
  ggplot(aes(Garage_Area, Sale_Price)) +
  geom_point() +
  labs(title = "Relation without outliers",
         x = "Size of garage",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

grid.arrange(Gar2, Gar1, ncol=2)

```

-   Here, we can say that we have some positive relation between sale price and Size of garage.

    -   But we can see there are some outliers in total basement size, which leads to the less weak relation.

-   The Correlation coefficient before removing outliers is **`r cor(ames['Garage_Area'], ames['Sale_Price'])`**

-   The Correlation coefficient after removing outliers is **`r cor(data2['Garage_Area'], data2['Sale_Price'])`**


## Relation between Sale Price and Total Basement size

```{r}

data3 <- ames %>% 
          filter((!Total_Bsmt_SF>3000)& (Total_Bsmt_SF > 0)) %>% 
          select(Total_Bsmt_SF,Sale_Price)

bsmt1 <- data3 %>% 
  ggplot(aes(Total_Bsmt_SF, Sale_Price)) +
  geom_point() +
  labs(title = "Relation without outliers",
         x = "Total Basement size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

bsmt2 <- ames %>% 
  ggplot(aes(Total_Bsmt_SF, Sale_Price)) +
  geom_point() +
  labs(title = "Relation with outliers",
         x = "Total Basement size",
         y = "Sale Price")+
  theme_replace() +
      scale_y_continuous(labels  = label_number(scale = 1e-05, prefix = "$", suffix = "hk", accuracy = 1)) +
  themes()

grid.arrange(bsmt2, bsmt1, ncol=2)

```

-   Here, we can say that we have some positive relation between sale price and total basement size.

    -   But we can see there are some outliers in total basement size, which leads to the less weak relation.

-   The Correlation coefficient before removing outliers is **`r cor(ames['Total_Bsmt_SF'], ames['Sale_Price'])`**

-   The Correlation coefficient after removing outliers is **`r cor(data3['Total_Bsmt_SF'], data3['Sale_Price'])`**

# Conclusion of EDA      
-   As we have seen so far that our three features have some outliers and by removing them we can get more strong relation which will help to improve performance of the model.

# Data Modeling          
- In this section we are going to learn how to fit a linear model for the each numerical feature we have seen above and how to interpret there equation.

## Model 1 (Sale Price and Above ground living area size)     

- When we fit the model, we can see the following values and with the help of that we can form an linear equation.

```{r}
m1 <- lm(Sale_Price ~ Gr_Liv_Area, data = data1)
tidy(m1)
```

$\hat{y}= 5905.812 + 116.874 \times Above-ground-living-area-size$        

- Y intercept value means that, For a house with 0 Above ground living area size, the house price will be **5905.812 USD**.   

- Slope of the fitted line means that, If we increase Above ground living area size by 1 Sqft, the price of the house increases with **116.874 USD**.

## Model 2 (Sale Price and Size of garage)

```{r}

m2 <- lm(Sale_Price ~ Garage_Area, data = data2)
tidy(m2)

```

$\hat{y}= 46925.28 + 277.37 \times Size-of-garage$

- Y intercept value means that, For a house with 0 Size of garage, the house price will be **46925.28 USD**.       

- Slope of the fitted line means that, If we increase Size of garage size by 1 Sqft, the price of the house increases with **277.37 USD**.

## Model 3 (Sale Price and Total Basement size)

```{r}
m3 <- lm(Sale_Price ~ Total_Bsmt_SF, data = data3)
tidy(m3)
```

$\hat{y}= 35393.94 + 137 \times Total-Basement-Size$
             
- Y intercept value means that, For a house with 0 total basement size , the house price will be **35393.94 USD**.       

- Slope of the fitted line means that, If we increase size of total basement by 1 Sqft, the price of the house with **137 USD**.



























